---
title: "Trial Report for 2 conditions"
author: "Gee"
date: "`r Sys.Date()`"
output: 
  html_document:
    css: styles.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,
  message = FALSE,
  fig.width = 6,
  fig.height = 5)
```

## Introduction
This report analyzes proteomics data. The analysis includes:

Data loading and quality control

Protein and peptide quantification

Statistical analysis

Visualization of results

```{r}
library(ggplot2)
library(dplyr)

custom_theme <- function() {
  theme_bw(base_size = 10) +
    theme(
      # Vertical x-axis labels (applies to ALL plots)
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
      
      # Unified text sizes
      axis.text = element_text(size = 10),
      axis.title = element_text(size = 12),
      plot.title = element_text(size = 14, face = "bold"),
      legend.text = element_text(size = 11),
      legend.title = element_text(size = 12),
      
      # Other consistent styling
      legend.position = "bottom",
      panel.grid.major.y = element_line(color = "gray90")
    )
}

# Make this theme apply to all plots automatically
theme_set(custom_theme())
```

## Load All Required Libraries
```{r Load All Required Libraries}
library(tidyverse)
library(ggrepel)
library(DEP)
library(missForest)
library(pheatmap)
library(openxlsx)
library(RColorBrewer)
library(ggpubr)
library(dichromat)
library(factoextra)
library(ggplotify)
library(OpenImageR)
library(svDialogs)
library(raster)
library(viridis)
library(patchwork)
```

## Data Loading and Preparation
```{r File Selection and Input Paths}
# Set working directory to script location
library(here)
sampleTypeDir <- file.path(dirname(rstudioapi::getSourceEditorContext()$path))
setwd(sampleTypeDir)
default_dir <- paste0(sampleTypeDir, "/*.*")

# Load custom functions
source("functions_ver2.3.R")

# Select input files interactively
protein_file_name <- choose.files(
  default = default_dir, 
  caption = "Please, select proteins file", 
  multi = FALSE
)
file_loc <- stringr::str_replace(protein_file_name, pattern = "_Proteins.txt", replacement = "")

sample_list_file_name <- choose.files(
  default = default_dir, 
  caption = "Please, select sample list file", 
  multi = FALSE
)

# Define peptides file
peptides_file_name <- paste0(file_loc, "_PeptideGroups.txt")
```

### Extract Project Name
```{r Extract Project name}
projectName <- t(data.frame(strsplit(protein_file_name, split = "\\\\")))
projectName <- projectName[,ncol(projectName)]
projectName <- stringr::str_replace(projectName, "_Proteins.txt", replacement = "")
```

### Create Output Directory
```{r create_output_directory, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
folderName <- "Results"
if (!dir.exists(file.path(sampleTypeDir, folderName))) {
  dir.create(file.path(sampleTypeDir, folderName))
}
plots_path <- file.path(sampleTypeDir, folderName, "/")
```

## Data Loading and Initial Processing
```{r Load Data files}
# Load protein data
proteinGroups <- read.delim(protein_file_name, check.names = FALSE)
peptides <- read.delim(peptides_file_name, check.names = FALSE)
sample_list <- read.delim(sample_list_file_name, check.names = FALSE)

# Keep only 2 groups for comparison
selected_groups <- c("Prototype", "MPSP")
sample_list <- sample_list %>%
  filter(Group %in% selected_groups)
```

### Set Filtering Criteria
```{r Set Filtering Criteria}
no_peptides <- 2
is_master_protein <- TRUE
```

### Prepare Protein Data
```{r prepare protein data}

### Prepare Protein Data

# Select columns for protein file 
list_of_columns_proteins <- c(
  "Number", 
  "Accession", 
  "Description", 
  "Master", 
  "Coverage [%]",
  "# Peptides", 
  "# PSMs", 
  "# Unique Peptides", 
  "MW [kDa]", 
  "Contaminant",
  "# Protein Groups", 
  "Sequence"
)

# Verify all requested columns exist
missing_cols <- setdiff(list_of_columns_proteins, names(proteinGroups))
if(length(missing_cols) > 0) {
  warning(paste("The following columns are missing:", paste(missing_cols, collapse=", ")))
}

# Select the identification columns
proteinGroups_identification <- proteinGroups %>%
  dplyr::select(any_of(list_of_columns_proteins), starts_with("Found in Sample:"))

# Select abundance columns - note your actual columns start with "Abundance:" not "Abundances:"
proteinGroups_abundances <- proteinGroups %>%
  dplyr::select(starts_with("Abundance:"))

# Handle empty values
proteinGroups_abundances[proteinGroups_abundances == ''] <- NA

# Select only samples from sample list
proteinGroups_abundances <- proteinGroups_abundances %>%
  dplyr::select(all_of(sample_list$RawFile))

# Update sample names
new_sample_names <- sample_list$Sample
colnames(proteinGroups_abundances) <- new_sample_names

# Merge identification and abundance data
proteinGroups_merged <- cbind(proteinGroups_identification, proteinGroups_abundances)

```

### Data Filtering
```{r Data Filtering}
# Apply filtering criteria
filtered <- proteinGroups_merged

if(is_master_protein == TRUE){
  filtered1 <- filtered[filtered$Master == "IsMasterProtein",]
} else {
  filtered1 <- filtered
}

filtered2 <- filtered1[filtered1$`# Peptides` >= no_peptides,]
filtered3 <- filtered2[filtered2$Contaminant != "True",]

proteinGroups_filtered <- filtered3
# Filter Peptides
peptides_filtered <- peptides %>%
  dplyr::select(Sequence, starts_with("Abundance:"), "XCorr (by Search Engine): Sequest HT") %>%
  filter(`XCorr (by Search Engine): Sequest HT` >= 1.5) %>%
  dplyr::select(Sequence, starts_with("Abundance:"))

# Rename columns to match sample names
colnames(peptides_filtered) <- c("Sequence", sample_list$Sample)
```

## Data Quality Assessment

### Protein and Peptide Quantification
```{r Protein and Peptide Quantification}
# Prepare extended color palette
extendet_pallete <- colorRampPalette(brewer.pal(8, name = "Dark2"))(
  length(unique(sample_list$Group))
)

# Protein quantification
proteinGroups_filtered_quality <- proteinGroups_filtered %>%
  dplyr::select(all_of(sample_list$Sample))

quantified_proteins <- data.frame(colSums(!is.na(proteinGroups_filtered_quality)))
quantified_proteins <- rownames_to_column(quantified_proteins)
colnames(quantified_proteins) <- c("Sample", "Quantified_Proteins")
quantified_proteins$Group <- sample_list$Group
```

## Visualization of Quantified Proteins

<details>
<summary>Click to expand: Quantified Proteins</summary>

```{r Visualization of Quantified Proteins}
if(nrow(sample_list) > 20){
  plot_quant_proteins <- ggplot(quantified_proteins, aes(x = Sample, y = Quantified_Proteins))+
    geom_point(aes(color = Group), size = 2)+
    geom_segment(aes(x = Sample, xend = Sample, y = 0, yend = Quantified_Proptides, color = Group))+
    theme_classic()+
    labs(y = "Number of filtered and quantified proteins", x = NULL)+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
    scale_color_manual(values = extendet_pallete)+
    geom_text(aes(label=Quantified_Proteins), vjust= -1.0, hjust = 0.5, angle = 90, 
              color="black", size=4)
} else {
  plot_quant_proteins <- ggplot(quantified_proteins, 
                               aes(x = Sample, y = Quantified_Proteins, fill = Group))+
    geom_bar(stat = "identity", position = "dodge")+
    geom_text(aes(label=Quantified_Proteins), vjust=0.5, hjust = 1.5, angle = 90, 
              color="white", size=4)+
    theme_classic()+
    labs(y = "Number of filtered and quantified proteins", x = NULL)+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
    scale_fill_manual(values = extendet_pallete)  
}

plot_quant_proteins
```
<details>

# Peptide Quantification
```{r Peptide Quantification}
peptides_filtered_quality <- peptides_filtered %>%
  dplyr::select(all_of(sample_list$Sample))

quantified_peptides <- data.frame(colSums(!is.na(peptides_filtered_quality)))
quantified_peptides <- rownames_to_column(quantified_peptides)
colnames(quantified_peptides) <- c("Sample", "quantified_peptides")
quantified_peptides$Group <- sample_list$Group
quantified_peptides$Group <- factor(quantified_peptides$Group, levels = unique(quantified_peptides$Group))
```
## Visualization of Quantified Peptides

<details>
<summary>Click to expand: Quantified Peptides</summary>

```{r Visualization of Quantified Peptides}
plot_quant_peptides <- ggplot(quantified_peptides, aes(x = Sample, y = quantified_peptides, fill = Group))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label=quantified_peptides), vjust=0.5, hjust = 1.5, angle = 90, color="white",
            position = position_dodge(0.9), size=4)+
  theme_classic()+
  labs(y = "Number of filtered and quantified peptides", x = NULL)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
  scale_fill_manual(values = extendet_pallete)

plot_quant_peptides
```
<details>

## QC – Average Quantified Peptides and Proteins

<details>
<summary>Click to expand: Average Quantified Peptides and Proteins</summary>

```{r Average Quantified Peptides and Proteins}
# Peptides

quant_peptides_average <- quantified_peptides %>%
  dplyr::group_by(Group) %>%
  dplyr::summarise(Mean = mean(quantified_peptides), SD = sd(quantified_peptides))

quant_peptides_average$Group <- factor(quant_peptides_average$Group, levels = unique(quant_peptides_average$Group))

p_peptides <- ggplot(quant_peptides_average, aes(x=Group, y=Mean, fill=Group)) + 
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar(aes(ymin=Mean-SD, ymax=Mean+SD), width=.2, position=position_dodge(.9)) +
  labs(y = "Average number of quantified peptides") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))+
  scale_fill_manual(values = extendet_pallete)

p_peptides

ggsave(filename = "plot_average_quantified_peptides.jpeg", plot = p_peptides, path = plots_path)

#Proteins

quant_proteins_average <- quantified_proteins %>%
  dplyr::group_by(Group) %>%
  dplyr::summarise(Mean = mean(Quantified_Proteins), 
                   SD = sd(Quantified_Proteins))

quant_proteins_average$Group <- factor(quant_proteins_average$Group, 
                                     levels = unique(quant_proteins_average$Group))

p_proteins <- ggplot(quant_proteins_average, aes(x=Group, y=Mean, fill=Group)) + 
  geom_bar(stat="identity", color="black", position=position_dodge(), width=0.7) +
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar(aes(ymin=Mean-SD, ymax=Mean+SD), width=.2, position=position_dodge(.9)) +
  labs(y = "Average number of quantified proteins") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))+
  scale_fill_manual(values = extendet_pallete)

p_proteins
```
<details>

```{r save plot_average_quantified_proteins}
ggsave(filename = "plot_average_quantified_proteins.jpeg", 
       plot = p_proteins, 
       path = plots_path,
       width = 8, 
       height = 6,
       dpi = 300)
```

## Re-order columns

```{r Re-order columns}
# Create Number column if it doesn't exist in BOTH data frames
if(!"Number" %in% names(proteinGroups_merged)) {
  proteinGroups_merged$Number <- seq_len(nrow(proteinGroups_merged))
}

if(!"Number" %in% names(proteinGroups_filtered)) {
  proteinGroups_filtered$Number <- seq_len(nrow(proteinGroups_filtered))
}

# Define column order
list_of_columns_proteins_order <- c(
  "Number", "Accession", "Description", "MW [kDa]",
  colnames(proteinGroups_abundances),
  "# Peptides", "# PSMs", "# Unique Peptides", "Coverage [%]", 
  "Contaminant", "# Protein Groups", "Master", "Sequence"
)

# Verify columns exist in BOTH data frames
existing_cols_merged <- intersect(list_of_columns_proteins_order, names(proteinGroups_merged))
existing_cols_filtered <- intersect(list_of_columns_proteins_order, names(proteinGroups_filtered))

# Get union of columns that exist in either data frame
final_cols <- union(existing_cols_merged, existing_cols_filtered)

# Apply column selection
proteinGroups_merged <- proteinGroups_merged %>%
  dplyr::select(all_of(final_cols))

proteinGroups_filtered <- proteinGroups_filtered %>%
  dplyr::select(all_of(final_cols))

# Warning about any missing columns
missing_cols <- setdiff(list_of_columns_proteins_order, final_cols)
if(length(missing_cols) > 0) {
  warning(paste("The following columns are missing from one or both data frames:",
                paste(missing_cols, collapse=", ")))
}
```
## QC – Histogram and Violin Plot of Intensities

<details>
<summary>Click to expand: Histogram and Violin Plot of Intensities</summary>

```{r qc_histogram_violin_plot, fig.width=6, fig.height=5}
# Prepare abundance data
qc_abundance <- proteinGroups_abundances %>%
  pivot_longer(cols = everything(), names_to = "Sample", values_to = "Abundance") %>%
  dplyr::mutate(
    Abundance = na_if(Abundance, 0), 
    Abundance_log = log2(Abundance)
  )

# Extract group from sample name
group_temp <- t(data.frame(str_split(qc_abundance$Sample, pattern = "_")))
qc_abundance$Group <- factor(group_temp[, 1], levels = unique(group_temp[, 1]))

# Median intensity for reference line
overall_median <- median(qc_abundance$Abundance_log, na.rm = TRUE)

# Histogram plot
hist_plot <- ggplot(qc_abundance, aes(x = Abundance_log)) +
  geom_histogram(bins = 30, fill = "steelblue", alpha = 0.8) +
  geom_vline(xintercept = overall_median, color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Protein Intensity Distribution",
    x = "Protein Intensity (log2)",
    y = "Count"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

# Violin plot
violin_plot <- ggplot(qc_abundance, aes(x = Group, y = Abundance_log, fill = Group)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  geom_boxplot(width = 0.1, fill = "white") +
  geom_hline(yintercept = overall_median, color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Protein Intensity by Group",
    x = "Group",
    y = "Protein Intensity (log2)"
  ) +
  scale_fill_manual(values = c(
  "Prototype" = "#E69F00",  
  "MPSP" = "#56B4E9"        
  )) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 12),
    legend.position = "none"
  )

# Show plots
hist_plot
violin_plot
```
<details>

```{r save plot}
ggsave(filename = "histogram_protein_abundance.jpeg", plot = hist_plot, path = plots_path)
ggsave(filename = "violinplot_protein_abundance.jpeg", plot = violin_plot, path = plots_path)
```

## Normalization, Filtering, and Imputation
```{r Normalization, Filtering, and Imputation}
# Normalization
protein_df <- proteinGroups_filtered %>%
  dplyr::select("Accession", all_of(sample_list$Sample))
rownames(protein_df) <- protein_df$Accession
protein_df <- protein_df[, -1]

df_normalized <- normalize_df(protein_df, 1) # 1 = median normalization

# Log2 transform
df_normalized <- log(df_normalized, 2)

# Filtering
group_info <- data.frame(Sample = sample_list$Sample, Group = sample_list$Group)
df_normalized_valid <- filter_valids_ml(df_normalized,
                                        conditions = unique(sample_list$Group),
                                        min_count = 2,
                                        at_least_one = TRUE)

# Imputation

input_matrix <- as.matrix(df_normalized_valid)

set.seed(123)  # ensures reproducibility
suppressMessages({
  missforest_result <- missForest(input_matrix, verbose = FALSE)
})

# Extract the imputed data
df_norm_valid_imp <- as.data.frame(missforest_result$ximp) 
save.image(file = paste0(Sys.Date(), "_my_work_space.RData"))
```
## Investigate Missing Values

<details>
<summary>Click to expand: Investigate Missing Values</summary>

```{r Missing values}
plotMissVal(proteinGroups_filtered_quality, plot_name = "non_valid")
MissValDist(proteinGroups_filtered)
```
<details>

## Imputation Effect – Histogram

<details>
<summary>Click to expand: Imputation Effect – Histogram</summary>

```{r Imputation effect-histogram}
# Prepare long-format data
df_norm_valid_imp_long <- df_norm_valid_imp %>%
  rownames_to_column("Protein") %>%
  pivot_longer(-Protein, names_to = "Sample", values_to = "Intensity")

# Plot
hist_imputation_effect<-ggplot(df_norm_valid_imp_long, aes(x = Intensity)) +
  geom_histogram(bins = 30, fill = "#00CED1", alpha = 0.7) +
  facet_wrap(~ Sample, scales = "free_y") +
  geom_vline(xintercept = median(df_norm_valid_imp_long$Intensity, na.rm = TRUE),
             color = "blue", linetype = "dashed", size = 1) +
  labs(
    title = "Protein Intensity Distribution",
    x = "Protein Intensity (log2)",
    y = "Count"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 12),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10)
  )

hist_imputation_effect
```
<details>

```{r save histogram imputation effect}
ggsave(filename = "hist_imputation_effect.jpeg", plot =hist_imputation_effect, path = plots_path)
```

## Scatter Plot Matrix

<details>
<summary>Click to expand: Multi Scatter Plot Matrix</summary>

```{r scatterplot matrix, fig.width=10, fig.height=10}
df_multi_scatter <- df_norm_valid_imp %>%
  dplyr::select(!contains("imputed")) %>%
  na.omit()

# Correlation panel
lower.panel <- function(x, y){
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- round(cor(x, y), digits=2)
  txt <- paste0("R = ", r)
  cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

# Customize upper panel
upper.panel <- function(x, y){
  points(x, y, pch = 19, size = 0.1, cex = 0.2)
}

pairs(df_multi_scatter[, 1:ncol(df_multi_scatter)], 
      lower.panel = lower.panel,
      upper.panel = upper.panel)
```
<details>

## Calculate CV- Coefficient of Variation

<details>
<summary>Click to expand: Coefficient of Variation- Violin Plot</summary>

```{r calculate CV-violin_cv_plot, fig.height=5, fig.width=6}
qc_cv <- inner_join(df_norm_valid_imp_long, group_info, by = "Sample")
qc_cv$IntensityUnlog <- 2^qc_cv$Intensity

qc_cv_df <- qc_cv %>%
  group_by(Protein, Group) %>%
  summarise(cv = cv(IntensityUnlog, na.rm = TRUE), .groups = "drop") %>%
  na.omit() %>%
  mutate(Group = factor(Group, levels = c("MPSP", "Other", "Prototype")))  # Ensure group order

# Enhanced plot
plot_violin_cv <- ggplot(qc_cv_df, aes(x = Group, y = cv, fill = Group)) +
  geom_violin(trim = FALSE, alpha = 0.7, width = 0.8, color = NA) +
  geom_boxplot(width = 0.15, fill = "white", color = "black", outlier.shape = NA) +
  scale_fill_manual(values = c(
  "MPSP" = "#8C8C8C",
  "Prototype" = "#C44E52"
)) +
  coord_cartesian(ylim = c(0, 50)) +  # Adjust based on your data range
  geom_hline(yintercept = 20, linetype = "dashed", color = "blue", linewidth = 1) +
  labs(
  x = NULL,
  y = "Coefficient of Variation (%)",
  title = "Technical Reproducibility Across Experimental Groups",
  caption = paste("n =", format(nrow(qc_cv_df)/3), "proteins per group | CV = SD/mean * 100")
) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.1, size = 14),
    axis.title.y = element_text(size = 14),
    panel.grid.major.y = element_line(color = "gray90")
  )

print(plot_violin_cv)
```
<details>

```{r save violin protein CV plot}
ggsave(filename = "violin_protein_cv.jpeg", plot = plot_violin_cv, path = plots_path)
```
## Protein CV Category Distribution 

<details>
<summary>Click to expand: Protein CV Category Distribution- Bar Plot</summary>
 
```{r create-cv-category-plot}
cv_barplot <- qc_cv_df %>%
  mutate(CV_Category = case_when(
    cv < 10 ~ "< 10%",
    cv >= 10 & cv < 20 ~ "10–20%", 
    cv >= 20 & cv < 30 ~ "20–30%",
    cv >= 30 & cv <= 100 ~ "30–100%",
    TRUE ~ "> 100%"  # optional fallback
  ))

# Reorder levels
cv_barplot$CV_Category <- factor(cv_barplot$CV_Category,
                                 levels = c("< 10%", "10–20%", "20–30%", "30–100%", "> 100%"))

# Summarize by group and CV bin
cv_barplot_summary <- cv_barplot %>%
  filter(CV_Category != "> 100%") %>%
  group_by(Group, CV_Category) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Group) %>%
  mutate(Percentage = Count / sum(Count) * 100)

# Define new elegant color palette
mycolors <- c(
  "< 10%"    = "#A6BDDB",
  "10–20%"   = "#74A9CF",
  "20–30%"   = "#2B8CBE",
  "30–100%"  = "#045A8D")

# Plot: Protein CV Category Distribution
barplot <- ggplot(cv_barplot_summary, aes(x = Group, y = Percentage, fill = CV_Category)) +
  geom_bar(stat = "identity", width = 0.85) +
  scale_fill_manual(values = mycolors) +
  labs(
    y = "Percentage of Proteins", 
    fill = "CV Category",
    title = "Protein CV Category Distribution"
  ) +
  geom_text(
    aes(label = sprintf("%.1f%%", Percentage)), 
    position = position_stack(vjust = 0.5),
    size = 4.5, color = "white"
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 14),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    legend.position = "right"
  )

barplot
```
<details>

```{r Protein CV Category Distribution - barplot }
ggsave(filename = "barplot_protein_cv_distribution.jpeg", plot = barplot, path = plots_path)
```


## QC: CV vs. Average Intensity (2D Density Plot)

<details>
<summary>Click to expand: Volcano Plot</summary>

```{r qc_cv_hexbin_density_plot, echo=TRUE, , fig.width=6, fig.height=5}
#Remove rows that are completely NA
df_norm_valid_imp <- df_norm_valid_imp[rowSums(is.na(df_norm_valid_imp)) < ncol(df_norm_valid_imp), ]

#Compute mean and CoV per protein
qc_stats <- df_norm_valid_imp %>%
  rowwise() %>%
  mutate(
    mean_intensity = mean(c_across(everything()), na.rm = TRUE),
    sd_intensity = sd(c_across(everything()), na.rm = TRUE),
    cv = ifelse(mean_intensity > 0, 100 * (sd_intensity / mean_intensity), NA_real_)
  ) %>%
  ungroup() %>%
  filter(is.finite(cv), is.finite(mean_intensity)) %>%
  filter(cv >= 0, cv <= 200)  # Optional threshold to clean noise

#Create Hexbin plot
hex_plot <- ggplot(qc_stats, aes(x = cv, y = mean_intensity)) +
  geom_hex(bins = 60) +
  scale_fill_viridis_c(option = "D") +
  labs(
    title = "2D Density Plot: CV vs Mean Intensity",
    x = "Coefficient of Variation (CV, %)",
    y = "Mean Intensity (log2)"
  ) +
    theme_minimal(base_size = 12) +
    theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 14),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    legend.position = "right"
  )
print(hex_plot)
```
<details>

```{r qc_cv_smoothed_density_plot, echo=TRUE, fig.height=6}
#Create Smoothed 2D Density plot
smoothed_density_plot <- ggplot(qc_stats, aes(x = cv, y = mean_intensity)) +
  stat_density_2d_filled(contour_var = "ndensity", bins = 30) +
  scale_fill_viridis_d(option = "D") +
  theme_minimal() +
  labs(
    title = "Smoothed Density: CV vs Mean Intensity",
    x = "Coefficient of Variation (CV, %)",
    y = "Mean Intensity (log2)"
  )
print(smoothed_density_plot)
```
## T-Test 
```{r t-test}
# Select only 2 groups of interest
selected_groups <- c("Prototype", "MPSP")

# Debug: rebuild group_info
group_info <- sample_list$Group
names(group_info) <- sample_list$Sample

# Filter for selected groups
group_filter <- group_info %in% selected_groups

# Subset the data
df_stat <- df_norm_valid_imp[, names(group_info)[group_filter]] %>%
  dplyr::select(!contains("imputed"))
group_info <- factor(group_info[group_filter], levels = selected_groups)

# Display table
table(group_info)

# T-test function
t.test.2 <- function(x, y) {
  tryCatch({
    if (length(unique(y)) != 2 || any(table(y) < 2)) {
      return(NA)
    }
    t.test(x ~ y)$p.value
  }, error = function(e) {
    return(NA)
  })
}

# Calculate p-values
p.values <- apply(df_stat , 1, t.test.2, group_info)
p.values.adj <- p.adjust(p.values, "BH")
proteins_stat <- cbind(df_stat, p.values, p.values.adj) %>% data.frame() %>%
  rownames_to_column(var = "Accession")
```
## Create df_description if missing
```{r Create Gene Description}
if(!exists("df_description")) {
  gene_name <- proteinGroups_filtered$Description
  gene_name <- str_split(gene_name, pattern = "GN=")
  collected_gene_names <- sapply(gene_name, function(x) str_split(x[2], pattern = " ")[[1]][1])
  
  gene_name <- proteinGroups_filtered$Description
  gene_name <- str_split(gene_name, pattern = " OS=")
  collected_description <- sapply(gene_name, function(x) x[1])
  
  df_description <- data.frame(
    "Accession" = proteinGroups_filtered$Accession,
    "Gene" = collected_gene_names,
    "ShortDesc" = collected_description
  )
}
```

## Volcano Plot
<details>
<summary>Click to expand: Volcano Plot</summary>

```{r volcano_plot, fig.width=6, fig.height=5}
# Ensure df_description exists
if(!exists("df_description")) stop("df_description not found - please run gene description section")
GroupA <- "Prototype"
GroupB <- "MPSP"
ratio_name <- paste0(GroupA, "/", GroupB)

df_volcano <- calculate_ratios_volcano_p_adjust(df_norm_valid_imp)

# Define significance
df_volcano$Significance <- ifelse(
  df_volcano$p.values.adj < 0.05 & abs(df_volcano$RatioLog) > 1,
  "Significant", "Not Significant"
)

volcano_plot <- ggplot(df_volcano, aes(x = RatioLog, y = -log10(p.values), col = Significance, label = TheLabel))+
  geom_point(size = 0.5)+
  theme_classic() +
  geom_vline(xintercept=c(-1, 1), col="blue", linetype = "longdash", size = 0.1) +
  geom_hline(yintercept = median(-log10(df_volcano$p.values), na.rm = TRUE), 
           color = "blue", linetype = "dashed", size = 1) +
  # Label top hits
  ggrepel::geom_text_repel(
    data = subset(df_volcano, Significance == "Significant" & -log10(p.values) > 2),
    size = 3,
    box.padding = 0.25,
    max.overlaps = 10
  ) +
  
  scale_color_manual(values = c("Not Significant" = "gray80", "Significant" = "#D73027")) +
  labs(
    title = "Volcano Plot: Prototype vs MPSP",
    x = paste0("Log2(", ratio_name, ")"),
    y = "-log10(p-value)"
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.position = "none"
  )

volcano_plot
```
</details>



```{r save volcano plot}
ggsave(filename = "volcano_plot.jpeg", plot = volcano_plot, dpi = 1200, path = plots_path)
ggsave(filename = "volcano_plot.svg", plot = volcano_plot, dpi = 1200, path = plots_path)
```

## Export Results
```{r}
openxlsx::write.xlsx(x = df_volcano, file = "proteins_df_stat.xlsx")
save.image(file = paste0(Sys.Date(), "_dummy_report_2.RData"))
```

## Summary

This report documents the complete data analysis pipeline for the mass spectrometry dataset, from data loading and filtering to normalization, imputation, statistical testing, and visualization.

Report generated on r Sys.Date() using RMarkdown.
